R-LIME\@: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-06-27 paper.tex
fix: remove warnings by compiler and linter

\def\mode{1} % 0: Graduation thesis, 1: ICPR paper

\ifnum\mode=0
  \documentclass[11pt]{article}
  \usepackage{luatexja}
  \usepackage{src/style/prmlthesis}
  \usepackage[a4paper]{geometry}
\else
  \documentclass[runningheads]{llncs}
\fi

@@

% The following code is used to remove the warning with bibliography:
% ------------------------------
% Underfull \hbox (badness 2653) in paragraph at lines 28--34
% ------------------------------
% cf. https://tex.stackexchange.com/questions/10924/underfull-hbox-in-bibliography
\apptocmd{\sloppy}{\hbadness 10000\relax}{}{}

% set the path for figures
\graphicspath{{./src/}{../python/data/}{src/experiments/}}

@@

% settings for cleveref
\crefname{table}{Table}{Tables}
\crefname{algorithm}{Algorithm}{Algorithms}
\ifnum\mode=0
  \crefname{figure}{Figure}{Figures}
\else
  \crefname{figure}{Fig.}{Figs.}
\fi

@@

\ifnum\mode=0
  \title{%
    R-LIME: LIME法における矩形制約と最適化
  }
  \etitle{%
    R-LIME: Rectangular Constraints and Optimization for Local Interpretable
    Model-agnostic Explanation Methods
  }
  \author{大原 玄嗣 (Genji Ohara)}
  \institute{%
    北海道大学 工学部 \\
    情報エレクトロニクス学科 \\
    情報理工学コース \\
    情報認識学研究室
  }
  \date{2024年3月}
  % \institute{%
  %   Laboratory for Pattern Recognition and Machine Learning,\\
  %   Division of Computer Science and Information Technology,\\
  %   Department of Electronics and Information Engineering,\\
  %   School of Engineering, Hokkaido University
  % }
\else
  % The title of the paper
  \title{%
    R-LIME: Rectangular Constraints and Optimization for Local Interpretable
    Model-agnostic Explanation Methods
  }

  % Abbreviated title for the running head
  \titlerunning{%
    R-LIME: Rectangular Constraints and Optimization for LIME Methods
  }

  % Author names and ORCID IDs
  \author{Genji Ohara\orcidID{0009-0000-5854-2820} \and
    Keigo Kimura\orcidID{0000-0002-3614-6568} \and
    Mineichi Kudo\orcidID{0000-0003-1013-3870}
  }

@@

  % The authors' affiliations
  \institute{Division of Computer Science and Information Technology\\
    Graduate School of Information Sci.\ and Tech.,
    Hokkaido University\\
    Sapporo 060-0814, JAPAN,\\
    \email{\{genji-ohara, kimura5, mine\}@ist.hokudai.ac.jp}}
\fi

@@

\ifnum\mode=0
  \pagestyle{plain}
  \pagenumbering{roman}
  \tableofcontents
  \newpage
  \listoffigures\listoftables\clearpage
  \AddToHook{cmd/section/before}{\clearpage}
  \pagestyle{plain}
  \pagenumbering{arabic}
\else
  % The abstract should briefly summarize the contents of the paper in
  % 150--250 words.
  \begin{abstract}
    In recent years,
    complex machine learning models have been introduced
    in various industrial fields due to their high accuracy.
    However,
    their increasing complexity has been a major obstacle to implementation
    in sensitive decision-making situations.
    In order to address this problem,
    various post-hoc explanation methods have been proposed,
    but they have not been able to achieve interpretability of
    both the explanation and its scope.
    We propose a new method, R-LIME,
    which interprets a complex classifier in an interpretable scope.
    R-LIME locally and linearly approximates a complex decision boundary
    of a black-box classifier in a rectangular region
    and maximizes the region as long as the approximation accuracy
    exceeds a given threshold.
    The resulting rectangular region is interpretable for users because it is
    expressed as a conjunction of feature predicates.
    Through qualitative and quantitative comparisons with the existing method
    on a real-world dataset,
    we demonstrate that R-LIME provides more reliable and interpretable
    explanations than existing methods.
    % 153 words
    \keywords{Interpretable machine learning \and Local surrogate model}
  \end{abstract}
\fi

\section{Introduction}
In recent years,
complex machine learning models, such as deep neural networks and random forests,
have been widely introduced in various industrial fields
due to their significant improvement in accuracy.
However,
their increasing complexity and black-box nature pose challenges,
particularly in critical decision-making scenarios such as healthcare and finance,
where the opacity of decision process becomes a major obstacle to implementation.
In order to address this problem,
there has been extensive research in the field of post-hoc explanations
for machine learning models~\cite{%
  guidotti2018local,radulovic2023bella,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure,
and the latter are further classified into \emph{global} and \emph{local} methods
based on their locality in input space~\cite{samek2021explaining}.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \ifnum\mode=0
      \vspace{-1.79cm}
    \else
      \vspace{-2.24cm}
    \fi
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method)
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

In this paper,
we focus on \emph{local} and \emph{model-agnostic} methods.
LIME~\cite{ribeiro2016why} and Anchor~\cite{ribeiro2018anchors}
are representative local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME linearly approximates the complex decision boundary
around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate its effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For this example,
users may apply the derived insights
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is obviously incorrect.
Anchor maximizes the coverage of a rectangular region containing the focal point
as long as the probability of the black-box classifier outputting
the same label as the focal point within the region exceeds a given threshold.
While Anchor provides an effective scope of the explanation,
users can get less insight compared to LIME\@.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that replacing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

@@

Furthermore, model-agnostic methods are categorized into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as partial dependence plots (PDP)~\cite{friedman2001greedy}
and accumulated local effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as local interpretable model-agnostic explanations (LIME)
\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
explain model's behavior in the vicinity of a specific input.
While they offer explanations more simple and accurate
than global methods,
the scope of the explanation is limited locally.

\section{Proposed Method}
\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption[Visual comparison of LIME, Anchor and R-LIME]{%
    Visual comparison of LIME, Anchor and R-LIME (our method).
    The dashed line represents the local linear approximation model,
    and the solid line represents the rectangular region containing the focal point.
  }
\end{figure}

@@

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)
  \cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation is performed by the following steps:
\begin{enumerate}
  \item Generating a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) $x$ is converted into a binary vector
        $x'\in\{0,1\}^{m'}$,
        (ii) perturbed samples are generated by drawing non-zero elements
        from $x'$ uniformly at random,
        and (iii) the perturbed samples are converted back to the original space.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically defined using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, it does not explicitly indicate the region for generating perturbed samples,
making it difficult for users to assess the effective scope of the explanation
\cite{ribeiro2018anchors}.

\subsubsection[Anchor]{%
  Anchor~\cite{ribeiro2018anchors}
}\label{sec:anchor}
Anchor maximizes the coverage of a rectangular region
containing the focal point $x$,
expressed as a conjunction of feature predicates (a rule)
as long as the probability of the black-box classifier $f$
outputting $f(x)$ within the region exceeds a given threshold $\tau$
(\cref{fig:anchor}).
It aims to highlight important features
contributing significantly to the output.
For a discrete $m$-dimensional input space $\ispace$,
a trained black-box classifier $f: \ispace \to \{0,1\}$,
an instance $x \in \ispace$
and a distribution $\mathcal{D}$ over the input space,
a rule $A(z) = a_{i_1}(z) \wedge a_{i_2}(z) \wedge \dots \wedge a_{i_t}(z)$ is defined.
The predicate $a_i(z)$ evaluates to true ($=1$) when $z_i = x_i$ and false ($=0$) otherwise.
The reliability of the explanation is defined as the ``accuracy'' of the rule,
and the generality of the explanation is defined as the ``coverage'' of the rule.
The accuracy $\Prec(A)$ and coverage $\Cov(A)$ of the rule $A$ are defined as follows:
\begin{align}
  \Prec(A) & =\mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=f(x)}], \label{eq:anchor_def_prec} \\
  \Cov(A)  & =\mathbb{E}_{z\sim\mathcal{D}(z)}[A(z)],
\end{align}
where $\mathcal{D}(z|A)$ is the conditional distribution in the region
where the rule $A$ returns true.
$\Prec(A)$ represents the probability that the output of $f$ matches
between the perturbation $z\sim\mathcal{D}(z|A)$ and the focal point $x$,
and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.
Anchor maximizes coverage as long as
the accuracy of the rule $A$ exceeds a given threshold $\tau$.
However, \cref{eq:anchor_def_prec} is not directly computable.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$,
the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const_prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main_problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

@@

Anchor maximizes the coverage of region $A$
as long as the probability of the output of the black-box classifier $f$ matching $f(x)$
within $A$ exceeds a given threshold $\tau$.
R-LIME, on the other hand, learns a linear classifier $g$
within the rectangular region $A$ and maximizes the coverage of $A$
under lower constraints on the accuracy of $g$.
We modify Anchor's definition of accuracy in \cref{eq:anchor_def_prec}
as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}], \label{eq:def_prec}
\end{equation}
where $G$ is a hypothesis space of possible linear classifiers.
By solving the optimization problem in \cref{eq:main_problem}
under the modified definition of accuracy in \cref{eq:def_prec},
we can select the rule that enables explanation with high accuracy and generality.

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{visual-rlime1}
    \includegraphics[width=0.32\textwidth]{visual-rlime2}
    \includegraphics[width=0.32\textwidth]{visual-rlime3}
    \caption[Overview of the R-LIME algorithm]{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$ is satisfied,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \def\myidt{\hspace{\algorithmicindent}}
  \ifnum\mode=0
    \begin{algorithm}[t]
      \small
      \input{src/tex/alg1}
    \end{algorithm}
  \else
    \begin{algorithm}[p]
      \small
      \input{src/tex/alg1}
    \end{algorithm}
  \fi

  \foreach \alg in {2,3,4}{%
      \begin{algorithm}[p]
        \small
        \input{src/tex/alg\alg}
      \end{algorithm}
    }
}
The algorithm of R-LIME
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
But greedy methods tend to converge to local optima,
so we use beam search, which selects multiple candidates at each iteration.
The pseudo-code is shown in \cref{alg:greedy_search}.

\subsubsection{Generating New Candidate Rules}
To generate new candidate rules,
one additional predicate is added to each of the $B$ candidate rules
selected in the previous iteration.
The pseudo-code is shown in \cref{alg:generate_cands}.
$T(x)$ is the set of predicates $\{a_1,\dots,a_m\}$,
where $a_i(z)$ evaluates to true when $z_i=x_i$ and false otherwise.
$T(x)\setminus A$ is the set of predicates in $T(x)$ not included in rule $A$.

\subsubsection{Searching Rules with Highest Accuracy}
Given the set of candidate rules $\cands$,
the algorithm selects the $B$ candidate rules with the highest accuracy.
This problem can be formulated
as best arm identification in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered as an arm,
and reward of arm $a_i$ follows a Bernoulli distribution
with $P(X=1)=\Prec(A_i)$.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using $z$ and $f(z)$ after each trial.
To efficiently search the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm \cite{kaufmann2013information}.
The pseudo-code is shown in \cref{alg:best_cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy the assumption.
This issue is discussed further in \cref{sec:reward}.

>>>>

\documentclass[runningheads]{llncs}

@@

% The following code is used to remove the warning with bibliography:
% ------------------------------
% Underfull \hbox (badness 2653) in paragraph at lines 28--34
% ------------------------------
% cf. https://tex.stackexchange.com/questions/10924/underfull-hbox-in-bibliography
\apptocmd{\sloppy}{\hbadness10000\relax}{}{}

% set the path for figures
\graphicspath{{./src/img}{./src/experiments/exp1}{./src/experiments/exp2}}

@@

% settings for cleveref
\crefname{table}{Table}{Tables}
\crefname{algorithm}{Algorithm}{Algorithms}
\crefname{figure}{Fig.}{Figs.}

@@

% The title of the paper
\title{%
  R-LIME\@: Rectangular Constraints and Optimization for Local Interpretable
  Model-agnostic Explanation Methods
}

% Abbreviated title for the running head
\titlerunning{%
  R-LIME\@: Rectangular Constraints and Optimization for LIME Methods
}

% Author names and ORCID IDs
\author{Genji Ohara\orcidID{0009-0000-5854-2820} \and % chktex 8
  Keigo Kimura\orcidID{0000-0002-3614-6568} \and      % chktex 8
  Mineichi Kudo\orcidID{0000-0003-1013-3870}          % chktex 8
}

@@

% The authors' affiliations
\institute{Division of Computer Science and Information Technology\\
  Graduate School of Information Sci.\ and Tech.,
  Hokkaido University\\
  Sapporo 060--0814, JAPAN,\\
  \email{\{genji-ohara, kimura5, mine\}@ist.hokudai.ac.jp}}

@@

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
\begin{abstract}
  In recent years,
  complex machine learning models have been introduced
  in various industrial fields due to their high accuracy.
  However,
  their increasing complexity has been a major obstacle to implementation
  in sensitive decision-making situations.
  In order to address this problem,
  various post-hoc explanation methods have been proposed,
  but they have not been able to achieve interpretability of
  both the explanation and its scope.
  We propose a new method, R-LIME,
  which interprets a complex classifier in an interpretable scope.
  R-LIME locally and linearly approximates a complex decision boundary
  of a black-box classifier in a rectangular region
  and maximizes the region as long as the approximation accuracy
  exceeds a given threshold.
  The resulting rectangular region is interpretable for users because it is
  expressed as a conjunction of feature predicates.
  Through qualitative and quantitative comparisons with the existing method
  on a real-world dataset,
  we demonstrate that R-LIME provides more reliable and interpretable
  explanations than existing methods.
  % 153 words
  \keywords{Interpretable machine learning \and Local surrogate model}
\end{abstract}

\section{Introduction}
In recent years,
complex machine learning models, such as deep neural networks and random forests,
have been widely introduced in various industrial fields
due to their significant improvement in accuracy.
However,
their increasing complexity and black-box nature pose challenges,
particularly in critical decision-making scenarios such as healthcare and finance,
where the opacity of decision process becomes a major obstacle to implementation.
In order to address this problem,
there has been extensive research in the field of post-hoc explanations
for machine learning models~\cite{%
  guidotti2018local,radulovic2023bella,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure,
and the latter are further classified into \emph{global} and \emph{local} methods
based on their locality in input space~\cite{samek2021explaining}.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \vspace{-2.24cm}
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method)
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

In this paper,
we focus on \emph{local} and \emph{model-agnostic} methods.
LIME~\cite{ribeiro2016why} and Anchor~\cite{ribeiro2018anchors}
are representative local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME linearly approximates the complex decision boundary
around the given focal point (\cref{fig:example-instance}),
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate its effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
~\cite{ribeiro2018anchors}.
For this example,
users may apply the derived insights
to the sentence ``This book is not good.'' % chktex 38
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is obviously incorrect.
Anchor maximizes the coverage of a rectangular region containing the focal point
as long as the probability of the black-box classifier outputting
the same label as the focal point within the region exceeds a given threshold.
While Anchor provides an effective scope of the explanation,
users can get less insight compared to LIME\@.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that replacing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME\@.

@@

Furthermore, model-agnostic methods are categorized into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as partial dependence plots (PDP)~\cite{friedman2001greedy}
and accumulated local effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as local interpretable model-agnostic explanations
(LIME)~\cite{ribeiro2016why} and Anchor~\cite{ribeiro2018anchors},
explain model's behavior in the vicinity of a specific input.
While they offer explanations more simple and accurate
than global methods,
the scope of the explanation is limited locally.

\section{Proposed Method}
\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-lime}
    \caption{%
      LIME\@: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-rlime3}
    \caption{%
      R-LIME\@:
      Maximizes coverage of a rectangular region
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption[Visual comparison of LIME, Anchor and R-LIME]{%
    Visual comparison of LIME, Anchor and R-LIME (our method).
    The dashed line represents the local linear approximation model,
    and the solid line represents the rectangular region containing the focal point.
  }
\end{figure}

@@

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)~\cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation is performed by the following steps:
\begin{enumerate}
  \item Generating a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) $x$ is converted into a binary vector
        $x'\in{\{0,1\}}^{m'}$,
        (ii) perturbed samples are generated by drawing non-zero elements
        from $x'$ uniformly at random,
        and (iii) the perturbed samples are converted back to the original space.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically defined using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, it does not explicitly indicate the region for generating perturbed samples,
making it difficult for users to assess the effective scope of the explanation
~\cite{ribeiro2018anchors}.

\subsubsection[Anchor]{Anchor~\cite{ribeiro2018anchors}}
Anchor maximizes the coverage of a rectangular region
containing the focal point $x$,
expressed as a conjunction of feature predicates (a rule)
as long as the probability of the black-box classifier $f$
outputting $f(x)$ within the region exceeds a given threshold $\tau$
(\cref{fig:anchor}).
It aims to highlight important features
contributing significantly to the output.
For a discrete $m$-dimensional input space $\ispace$,
a trained black-box classifier $f: \ispace \to \{0,1\}$,
an instance $x \in \ispace$
and a distribution $\mathcal{D}$ over the input space,
a rule $A(z) = a_{i_1}(z) \wedge a_{i_2}(z) \wedge \dots \wedge a_{i_t}(z)$ is defined.
The predicate $a_i(z)$ evaluates to true ($=1$) when $z_i = x_i$ and false ($=0$) otherwise.
The reliability of the explanation is defined as the ``accuracy'' of the rule,
and the generality of the explanation is defined as the ``coverage'' of the rule.
The accuracy $\Prec(A)$ and coverage $\Cov(A)$ of the rule $A$ are defined as follows:
\begin{align}
  \Prec(A) & =\mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=f(x)}], \label{eq:anchor-def-prec} \\
  \Cov(A)  & =\mathbb{E}_{z\sim\mathcal{D}(z)}[A(z)],
\end{align}
where $\mathcal{D}(z|A)$ is the conditional distribution in the region
where the rule $A$ returns true.
$\Prec(A)$ represents the probability that the output of $f$ matches
between the perturbation $z\sim\mathcal{D}(z|A)$ and the focal point $x$,
and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.
Anchor maximizes coverage as long as
the accuracy of the rule $A$ exceeds a given threshold $\tau$.
However, \cref{eq:anchor-def-prec} is not directly computable.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$,
the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const-prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main-problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

@@

Anchor maximizes the coverage of region $A$
as long as the probability of the output of the black-box classifier $f$ matching $f(x)$
within $A$ exceeds a given threshold $\tau$.
R-LIME, on the other hand, learns a linear classifier $g$
within the rectangular region $A$ and maximizes the coverage of $A$
under lower constraints on the accuracy of $g$.
We modify Anchor's definition of accuracy in \cref{eq:anchor-def-prec}
as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}], \label{eq:def-prec}
\end{equation}
where $G$ is a hypothesis space of possible linear classifiers.
By solving the optimization problem in \cref{eq:main-problem}
under the modified definition of accuracy in \cref{eq:def-prec},
we can select the rule that enables explanation with high accuracy and generality.

\subsection{Algorithm}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{visual-rlime1}
    \includegraphics[width=0.32\textwidth]{visual-rlime2}
    \includegraphics[width=0.32\textwidth]{visual-rlime3}
    \caption[Overview of the R-LIME algorithm]{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$ is satisfied,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \def\myidt{\hspace{\algorithmicindent}}
  \begin{algorithm}[p]
    \small
    \input{src/tex/alg1}
  \end{algorithm}

  \begin{algorithm}[p]
    \small
    \input{src/tex/alg2}
  \end{algorithm}

  \begin{algorithm}[p]
    \small
    \input{src/tex/alg3}
  \end{algorithm}

  \begin{algorithm}[p]
    \small
    \input{src/tex/alg4}
  \end{algorithm}
}
The algorithm of R-LIME
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main-problem},
greedy search are often used.
But greedy methods tend to converge to local optima,
so we use beam search, which selects multiple candidates at each iteration.
The pseudo-code is shown in \cref{alg:greedy-search}.

\subsubsection{Generating New Candidate Rules}
To generate new candidate rules,
one additional predicate is added to each of the $B$ candidate rules
selected in the previous iteration.
The pseudo-code is shown in \cref{alg:generate-cands}.
$T(x)$ is the set of predicates $\{a_1,\dots,a_m\}$,
where $a_i(z)$ evaluates to true when $z_i=x_i$ and false otherwise.
$T(x)\setminus A$ is the set of predicates in $T(x)$ not included in rule $A$.

\subsubsection{Searching Rules with Highest Accuracy}
Given the set of candidate rules $\cands$,
the algorithm selects the $B$ candidate rules with the highest accuracy.
This problem can be formulated
as best arm identification in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered as an arm,
and reward of arm $a_i$ follows a Bernoulli distribution
with $P(X=1)=\Prec(A_i)$.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using $z$ and $f(z)$ after each trial.
To efficiently search the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm~\cite{kaufmann2013information}.
The pseudo-code is shown in \cref{alg:best-cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However,
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged,
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy the assumption.
This issue is discussed further in \cref{sec:reward}.
